Hydrodynamics and rheology of a vesicle doublet suspension

### 2019-06-21 vesicle_code/docs/adhesion/adhesion.tex
yny

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
%\documentclass[aps,prb,showpacs,amsmath,amssymb,superscriptaddress]{revtex4-1}
\documentclass[prf,superscriptaddress,showkeys]{revtex4-1}

@@

\usepackage{amsmath,amssymb}
\usepackage[titletoc]{appendix}
\usepackage{todonotes}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{stmaryrd}
%\usepackage{subfig}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\newcommand{\subfigimg}[3][,]{%
  \setbox1=\hbox{\includegraphics[#1]{#3}}% Store image in box
  \leavevmode\rlap{\usebox1}% Print image
  \rlap{\hspace*{0pt}\raisebox{\dimexpr\ht1-0\baselineskip}{\bf
  \normalsize #2}}% Print label
  \phantom{\usebox1}% Insert appropriate spacing
}

@@

\begin{abstract}
The dynamics of an adhesive two-dimensional vesicle doublet under
various flow conditions is investigated numerically using a high-order,
adaptive-in-time boundary integral method. In a quiescent flow, two
nearby vesicles move slowly towards each other under the adhesive
potential, pushing out fluid between them to form a vesicle doublet at
equilibrium. A lubrication analysis on such draining of a thin film
gives the dependencies of draining time on adhesion strength and
separation distance that are in good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is
dynamically placed in the middle of a vesicle doublet through an active
control loop, novel dynamics of a vesicle doublet is observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the diverging stream,
thus providing a simple method to measure the adhesion strength between
two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form provided that the adhesion strength is sufficiently
large at a given vesicle reduced area. Once a doublet is formed, the
effective shear viscosity of a dilute suspension of vesicle doublets is
found to be a function of  reduced area.  Results from these numerical
studies and analysis shed light on the hydrodynamic and rheological
consequences of adhesive interactions between vesicles in a viscous
fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-treading dynamics transitions to tumbling
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in the effective shear viscosity of the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{Vitkova2008_BJ,DeschampsKantsler2009_PNAS,
KantslerSegreSteinberg2008_EPL, ZabuskySegreDeschamps2011_PoF}.  Between
tank-treading and tumbling vesicle dynamics, a breathing (tremble) mode
is also observed~\cite{Misbah2006_PRL, KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to alter the effective
shear viscosity.  In an extensional flow (planar or uniaxial), vesicle
shape dynamics depends sensitively on the vesicle reduced volume and the
elastic capillary number~\cite{KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2013_JFM, Narsimhan2014_JFM,
DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape and oscillatory
undulation of the vesicle membrane are two examples of the complex
vesicle hydrodynamics in an extensional flow.

@@

\begin{figure}[htp]
  \includegraphics[height=0.26\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
  \includegraphics[height=0.26\textwidth]{figs/adR4em1adS7em1_finalInclinationAngle.pdf}
  \includegraphics[height=0.26\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \caption{\label{fig:extensionalInclinationAngle} {\em Left}: The
  inclination angle of a doublet formed by vesicles of reduced area
  $\Delta A = 0.7$ in an extensional flow with an extensional rate
  $\chi$.  {\em Middle}: The final inclination angle of doublets.  At
  larger extensional rates, the doublet is broken by the flow.  {\em
  Right}: A phase diagram of the long-time behavior of a fluid trap
  formed by two adhering vesicles in an extensional flow.  Results are
  not reported for some of the low extensional rates since the doublet
  had not tilted to an equilibrium inclination angle for the given time
  horizon.}
\end{figure}

Finally, to further characterize the rotating dynamics of a doublet in
the fluid trap, we compute and plot the vorticity and total force on a
doublet with identical vesicles with reduced area $\Delta A=0.7$,
Hamaker constant $\mathcal{H} = 0.7$, separation distance $\delta =
0.4$, and extensional rate $\chi = 0.07$ in
Figure~\ref{fig:extensionalVorticity}.  Starting from an initial inclination angle $\theta=\pi$, the doublet 
starts to rotate around $t\sim 250$.  We notice that the net force on
each vesicle is almost anti-parallel to the diverging flow in the
far-field so to balance the hydrodynamic drag force that pulls the
vesicles apart.  Then, as the doublet starts to rotate, each vesicle
experiences a drag force from the diverging streaming
flows.  
%
For small extensional rate, the rotation continues until
$\theta = \pi/2$ so the doublet remains perfectly symmetric with respect
to their mid-plane, and each vesicle experiences drag forces (from the diverging flow) that cancel
each other.  
%
For slightly higher extensional rate in Figure~\ref{fig:extensionalVorticity}, we observe
that the two vesicles slide against each other (around $t=280$) as the doublet rotates.
The slight sliding motion leads to a transition in the contact region
membrane shape from a bulging profile to an S-shape profile of nearly
uniform height except near the edge of contact region.  This adjustment
greatly reduces the net force on each vesicle, leading to an equilibrium
configuration with an inclination angle $\theta > \pi/2$ as shown in the
left of Figure~\ref{fig:extensionalInclinationAngle}.  Finally, the
vorticity and the flow between the vesicles approaches zero and the
result is a fluid trap containing the doublet.
\begin{figure}[htp]
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity01.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity02.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity03.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity04.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity05.png} \\
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity06.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity07.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity08.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity09.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity10.png}
  \caption{\label{fig:extensionalVorticity} The vorticity and total
  force of two vesicles in a doublet.  The reduced area is $\Delta A =
  0.7$, the separation distance is $\delta = 0.4$, the Hamaker constant
  is $\mathcal{H} = 0.7$, and the extensional rate is $\chi = 0.07$.}
\end{figure}

@@

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
separation distance to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In the left plot of
Figure~\ref{fig:sflow_distance}, we plot the minimum distance
between the vesicles as a function of time for various Hamaker
constants.  The red curves correspond to Hamaker constants that are too
small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.  Figure~\ref{fig:sflow_phase_diagram} indicates whether a
doublet with reduced area $\Delta A = 0.9$ is formed (blue) or not (red)
for a variety of shear rates, Hamaker constants, and at two different
separation distances.  The plot includes a line of best fit between the
stable and unstable regions.  Since the transition is linear, this
indicates that a dimensionless number involving the ratio of the Hamaker
constant could be used to determine if a doublet forms or not.

\begin{figure}
  \includegraphics[height=0.27\textwidth]{figs/shear_adR4em1Chi5em1_ra090.pdf}
  \caption{\label{fig:sflow_distance} {\em Left}: The distance between a
  pair of vesicles in a planar shear flow with shear rate $\chi=0.5$,
  and separation distance $\delta = 0.4$.  The different lines
  correspond to a linear spacing of Hamaker constants ranging from
  $\mathcal{H}=0.1$ and $\mathcal{H}=1.0$.  {\em Right}:
  \todo[inline]{frequency of the oscillation}}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[height=0.37\textwidth]{figs/shear_adR1em1_ra090_phaseDiagram.pdf}
  \includegraphics[height=0.37\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{\label{fig:sflow_phase_diagram} Phase diagram of vesicle
  hydrodynamics in a shear flow with $\delta = 0.1$ (left) and $\delta =
  0.4$ (right) with reduced area $\Delta A = 0.9$. The critical Hamaker
  constant $\mathcal{H}$ for binding of two vesicles in a shear flow
  depends on the shear rate $\chi$.  For a doublet to form, lower
  separation distances require a stronger Hamaker constants at a given
  shear rate.  The stability limit appears to be linear, indicating that
  there is potentially a stability criteria that depends on a
  dimensionless number involving the ratio of the Hamaker constant and
  shear rate.}
\end{figure}

Finally, we investigate the rheology of a suspension of a doublet by
computing the effective viscosity of a doublet and compare it to the
effective viscosity of a single tank-treading vesicle. The effective
viscosity is defined as the viscosity of a homogeneous Newtonian fluid
with the same energy dissipation per macroscopic element of fluid.  In a
simple shear flow, the intrinsic viscosity, $[\mu]$ is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 (T_e - T_i)} \int_{T_i}^{T_e} 
  \langle \tau_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds,
\end{align*}
$\phi$ is the area fraction of vesicles, $\tau$ is the stress due to the
vesicles, $\langle \cdot \rangle$ is the spatial average, $|\omega|$ is
the total area of both vesicles, and $\xxi$ is the traction as defined
in equation~\eqref{eqn:traction}.  In
Figure~\ref{fig:shearIntrinsicViscosity}(a), we compare the intrinsic
viscosity of a single tank-treading vesicle to a doublet.  To validate
our simulations, we superimpose (black marks) the intrinsic viscosity
calculated by Ghigliotti {\em et
al.}~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
slightly increases the intrinsic viscosity at all the reduced areas.  To
further characterize the effect of adhesion, in
Figure~\ref{fig:shearIntrinsicViscosity}(b) we decompose the intrinsic
viscosity into the contributions from the bending and tension (blue),
and the contribution from the adhesion (red).  We also superimpose the
intrinsic viscosity of a single tank-treading vesicle (dashed curve) to
demonstrate that the bending and tension of the doublet behave
similarly, but not identically, to a dilute suspensions of non-adhering
tank-treading vesicles.  We see that the effect of the adhesion on the
intrinsic viscosity is largest for vesicles with small reduced areas.

\begin{figure}[]
  \begin{tabular}{@{}p{0.45\linewidth}@{\quad}p{0.45\linewidth}@{}}
  \subfigimg[width=\linewidth]{(a)}{figs/doublet_decomp.pdf} &
  \subfigimg[width=\linewidth]{(b)}{figs/doublet_decomp.pdf}
  \end{tabular}
%  \centering
%  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
%  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} (a) The intrinsic
  viscosity of a tank-treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the separation distance is $\delta = 0.4$.  The black marks
  denote intrinsic viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  (b) The
  decomposition of the intrinsic viscosity into the contributions for
  the bending and tension (blue) of the vesicles, and the contribution
  from the vesicle adhesion (red).  Also included is the intrinsic
  viscosity of a single tank-treading vesicle (dashed black line).}
\end{figure}

>>>>

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
%\documentclass[aps,prb,showpacs,amsmath,amssymb,superscriptaddress]{revtex4-1}
\documentclass[prf,superscriptaddress,showkeys,longbibliography]{revtex4-1}

@@

\usepackage{amsmath,amssymb}
\usepackage[titletoc]{appendix}
\usepackage{todonotes}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{stmaryrd}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

@@

\begin{abstract}
The dynamics of an adhesive vesicle doublet under various flow conditions is
investigated numerically using a high-order, adaptive-in-time boundary
integral method. In a quiescent flow, two nearby vesicles
move slowly towards each other under the adhesive potential, pushing out
fluid between them to form a vesicle doublet at equilibrium. A lubrication
analysis on such draining of a thin film gives the dependencies of
draining time on adhesion strength and separation distance that are in
good agreement with numerical results.
%
In a microfluid trap where the stagnation of an extensional flow is dynamically
placed in the middle of a vesicle doublet through an active control
loop, novel dynamics of a vesicle doublet is observed. Numerical
simulations show that there exists a critical extensional flow rate
above which adhesive interaction is overcome by the 
diverging stream, thus providing a simple method to measure the adhesion
strength between two vesicle membranes. 
%
In a planar shear flow, numerical simulations reveal that a vesicle
doublet may form provided that the adhesion
strength is sufficiently large at a given vesicle reduced area. Once a
doublet is formed, the effective shear viscosity of a dilute suspension
of vesicle doublets is found to be a function of  reduced area.  Results from these numerical studies and
analysis shed light on the hydrodynamic and rheological consequences of adhesive interactions
between vesicles in a viscous fluid.
%\begin{description}
%\item[Usage]
%Secondary publications and information retrieval purposes.
%\item[PACS numbers]
%May be entered using the \verb+\pacs{#1}+ command.
%\item[Structure]
%You may use the \texttt{description} environment to structure your abstract;
%use the optional argument of the \verb+\item+ command to give the category of each item. 
%\end{description}
\end{abstract}

@@

Hydrodynamics of a single vesicle in Stokes flow has been extensively
investigated. In a planar shear flow, the vesicle hydrodynamics is
characterized by the reduced volume (reduced area in two dimensions),
viscosity contrast between interior and exterior fluids, and shear rate
of the imposed far-field fluid flow. In addition, a vesicle with a rigid
particle inside is also investigated as a biological mimic of a
eukaryotic cell with a nucleus that occupies nearly half of the
intracellular volume~\cite{Veerapaneni2011_PRL}. Small-deformation
analysis shows that a vesicle tank-treads in a planar shear flow for low
viscosity contrast and shear rate. At high viscosity contrast the
tank-treading dynamics transitions to tumbling
dynamics~\cite{Misbah2006_PRL, Vlahovska2007_PRE}, and this leads to a
transition in the effective shear viscosity of the vesicle
suspension~\cite{Misbah2006_PRL,Vitkova2008_BJ} that is also validated
by direct numerical simulations~\cite{GhigliottiBibenMisbah2010_JFM} and
experiments~\cite{DeschampsKantsler2009_PNAS,
KantslerSegreSteinberg2008_EPL, ZabuskySegreDeschamps2011_PoF}.  Between
tank-treading and tumbling vesicle dynamics, a breathing (tremble) mode
is also observed~\cite{Misbah2006_PRL, KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2011_JFM, SpannZhaoShaqfeh2014_PoF} to alter the effective
shear viscosity.  In an extensional flow (planar or uniaxial), vesicle
shape dynamics depends sensitively on the vesicle reduced volume and the
elastic capillary number~\cite{KantslerSegreSteinberg2008_PRL,
ZhaoShaqfeh2013_JFM, Narsimhan2014_JFM,
DahlNarsimhanGouveia2016_SoftMatt}: Asymmetric shape and oscillatory
undulation of the vesicle membrane are two examples of the complex
vesicle hydrodynamics in an extensional flow.

@@

\begin{figure}[htp]
  \includegraphics[height=0.27\textwidth]{figs/adR4em1adS7em1_ra070_inclinationAngle.pdf}
  \includegraphics[height=0.27\textwidth]{figs/adR4em1adS7em1_finalInclinationAngle.pdf}
  \includegraphics[height=0.27\textwidth]{figs/extensional_adR4em1adS7em1_phaseDiagram.pdf}
  \caption{\label{fig:extensionalInclinationAngle} {\em Left}: The
  inclination angle of a doublet formed by vesicles of reduced area
  $\Delta A = 0.7$ in an extensional flow with an extensional rate
  $\chi$.  {\em Middle}: The final inclination angle of doublets.  At
  larger extensional rates, the doublet is broken by the flow.  {\em
  Right}: A phase diagram of the long-time behavior of a fluid trap
  formed by two adhering vesicles in an extensional flow.  Results are
  not reported for some of the low extensional rates since the doublet
  had not tilted to an equilibrium inclination angle for the given time
  horizon.}
\end{figure}

Finally, to further characterize the rotating dynamics of a doublet in
the fluid trap, we compute and plot the vorticity and total force on a
doublet with identical vesicles with reduced area $\Delta A=0.7$,
Hamaker constant $\mathcal{H} = 0.7$, separation distance $\delta =
0.4$, and extensional rate $\chi = 0.7$ in
Figure~\ref{fig:extensionalVorticity}.  Starting from an initial inclination angle $\theta=\pi$, the doublet 
starts to rotate around $t\sim 250$.  We notice that the net force on
each vesicle is almost anti-parallel to the diverging flow in the
far-field so to balance the hydrodynamic drag force that pulls the
vesicles apart.  Then, as the doublet starts to rotate, each vesicle
experiences a drag force from the diverging streaming
flows.  
%
For small extensional rate, the rotation continues until
$\theta = \pi/2$ so the doublet remains perfectly symmetric with respect
to their mid-plane, and each vesicle experiences drag forces (from the diverging flow) that cancel
each other.  
%
For slightly higher extensional rate in Figure~\ref{fig:extensionalVorticity}, we observe
that the two vesicles slide against each other (around $t=280$) as the doublet rotates.
The slight sliding motion leads to a transition in the contact region
membrane shape from a bulging profile to an S-shape profile of nearly
uniform height except near the edge of contact region.  This adjustment
greatly reduces the net force on each vesicle, leading to an equilibrium
configuration with an inclination angle $\theta > \pi/2$ as shown in the
left of Figure~\ref{fig:extensionalInclinationAngle}.  Finally, the
vorticity and the flow between the vesicles approaches zero and the
result is a fluid trap containing the doublet.
\begin{figure}[htp]
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity01.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity02.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity03.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity04.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity05.png} \\
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity06.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity07.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity08.png}
  \includegraphics[height = 0.26\textwidth,trim={6cm 1cm 5cm
  0cm},clip]{figs/extensionalVorticity09.png}
  \includegraphics[height = 0.26\textwidth,trim={5cm 1cm 2cm
  0cm},clip]{figs/extensionalVorticity10.png}
  \caption{\label{fig:extensionalVorticity} The vorticity and total
  force of two vesicles in a doublet.  The reduced area is $\Delta A =
  0.7$, the separation distance is $\delta = 0.4$, the Hamaker constant
  is $\mathcal{H} = 0.7$, and the extensional rate is $\chi = 0.7$.}
\end{figure}

@@

We characterize the effect of the Hamaker constant and shear rate by
determining a critical Hamaker constant, $\mathcal{H}_c$, which
determines if the vesicles form a doublet or separate.  We fix the
separation distance to $\delta = 0.4$ and the vesicle reduced area to
$\Delta A = 0.9$.  In the left plot of
Figure~\ref{fig:sflow_phase_diagram}, we plot the minimum distance
between the vesicles as a function of time for various Hamaker
constants.  The red curves correspond to Hamaker constants that are too
small to bind the vesicles into a doublet, and the blue curves
correspond to Hamaker constants that are sufficiently large to form a
doublet.

\begin{figure}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1Chi1e0_ra-090.pdf}
  \includegraphics[height=0.35\textwidth]{figs/shear_adR4em1_ra090_phaseDiagram.pdf}
  \caption{\label{fig:sflow_phase_diagram} {\em Left}: Distance between
  a pair of vesicles in a planar shear flow with shear rate $\chi=0.5$,
  and separation distance $\delta = 0.4$.  The different lines
  correspond to a linear spacing of Hamaker constants ranging from
  $\mathcal{H}=0.1$ and $\mathcal{H}=1.0$.  {\em Right}: Phase diagram
  of vesicle hydrodynamics in a shear flow.  At a fixed reduced area
  $\Delta A=0.90$ and $\delta = 0.4$, the critical Hamaker constant
  $\mathcal{H}$ for binding of two vesicles in a shear flow depends on
  the shear rate $\chi$.}
\end{figure}

Finally, we investigate the rheology of a suspension of a doublet by
computing the effective viscosity of a doublet and compare it to the
effective viscosity of a single tank-treading vesicle. The effective
viscosity is defined as the viscosity of a homogeneous Newtonian fluid
with the same energy dissipation per macroscopic element of fluid.  In a
simple shear flow, the intrinsic viscosity, $[\mu]$ is
\begin{align*}
  [\mu]:= \frac{\mu_{\mathrm{eff}} - \mu_0}{\phi \mu_0} = 
  \frac{1}{\chi \mu_0 (T_e - T_i)} \int_{T_i}^{T_e} 
  \langle \tau_{12} \rangle dt,
\end{align*}
where
\begin{align*}
  \langle \tau \rangle = \frac{1}{|\omega|} \int_{\gamma}
    \xxi \otimes \xx ds,
\end{align*}
$\phi$ is the area fraction of vesicles, $\tau$ is the stress due to the
vesicles, $\langle \cdot \rangle$ is the spatial average, and $\xxi$ is
the traction as defined in equation~\eqref{eqn:traction}.  In the left
plot of Figure~\ref{fig:shearIntrinsicViscosity}, we compare the
intrinsic viscosity of a single tank-treading vesicle to a doublet.  To
validate our simulations, we superimpose (black marks) the intrinsic
viscosity calculated by Ghigliotti {\em et
al.}~\cite{GhigliottiBibenMisbah2010_JFM}.  The presence of the doublet
significantly increases the intrinsic viscosity at all the reduced
areas.  To further characterize the effect of adhesion, in the right
plot of Figure~\ref{fig:shearIntrinsicViscosity} we decompose the
intrinsic viscosity into the contributions from the bending and tension
(blue), and the contribution from the adhesion (red).  We also
superimpose twice the intrinsic viscosity of a single tank-treading
vesicle (dashed curve) to demonstrate that the bending and tension of
the doublet behave similarly, but not identically, to a dilute
suspensions of non-adhering tank-treading vesicles.  We see that the
effect of the adhesion on the intrinsic viscosity is largest for
vesicles with small reduced areas.

\begin{figure}[htp]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/shear2Ves_adR4em1adS7em1Chi5em1.pdf}
  \includegraphics[width=0.45\textwidth]{figs/doublet_decomp.pdf}
  \caption{\label{fig:shearIntrinsicViscosity} {\em Left}: The intrinsic
  viscosity of a tank-treading vesicle (blue) and a doublet (red).  The
  shear rate is $\chi = 0.5$, the Hamaker constant is $\mathcal{H} =
  0.7$, and the separation distance is $\delta = 0.4$.  The black marks
  denote intrinsic viscosity values computed by Ghigliotti {\em et
  al.}~\cite{GhigliottiBibenMisbah2010_JFM} (cf.~Figure 5).  {\em
  Right}: The decomposition of the intrinsic viscosity into the
  contributions for the bending and tension (blue) of the vesicles, and
  the contribution from the vesicle adhesion (red).  Also included is
  twice the intrinsic viscosity of a single tank-treading vesicle
  (dashed black line).}
\end{figure}
